\documentclass{llncs}
\input{header}

\usepackage[colorlinks=true]{hyperref}
\usepackage{graphicx}
\usepackage{listings}

\begin{document}

\title{\textsc{jCupid}: A dynamic analysis tool for detecting side channels in Java programs}
\author{
  Ian Martiny
  \and
  Eric Wustrow
  \and
  Pavol {\v C}ern\'y
  \and
  Ashutosh Trivedi
}
\institute{University of Colorado Boulder, USA \\
  \email{\{ian.martiny, ewust, pavol.cerny, ashutosh.trivedi\}@colorado.edu}}


\maketitle

\begin{abstract}
  Side-channel based vulnerabilities continue to plague modern security-critical
  applications.
  From leaking private-keys to exposing sensitive data, 
  protecting against these threats remains a difficult task for developers. 
  A common defense against side-channels is to write constant-time code.
  However, ensuring that a complex program has no data-dependent control flow
  can be a laborious challenge, and can on occasion introduce or enable new side
  channel vulnerabilities.
  We present \jcupid, a tool that aids developers in detecting and
  removing a particular class of side channels from their code.
  \jcupid dynamically tests Java programs for timing side channels by examining
  the sequence of bytecodes that are executed for multiple inputs of the same size.
  Our tool optimizes this process by first computing only hashes of bytecode sequences,
  so that comparing the full sequences becomes necessary only when a side-channel has been detected.
  Once two inputs are found that cause a different sequence of bytecodes to be
  executed, \jcupid can inform the developer of the offending line of source code
  responsible for the divergence.
  While our tool can naively fuzz an instrumented program, it can also be used
  in conjuction with a Concolic execution testing engine.
  We show that instrumenting programs for use with \jcupid is simple for
  developers, and that our tool can reveal timing side-channel vulnerabilities
  that are not immediately obvious from source code inspection.
\end{abstract}



\section{Introduction}
\label{sec:introduction}
\input{introduction}


\section{\textsc{jCupid}}
\label{sec:jcupid}
\input{jcupid}

\section{Evaluation}
\label{sec:evaluation}
\input{evaluation}

\section{Future Work}
\label{sec:future}
\input{future}

\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}


\bibliographystyle{plain}
\bibliography{papers}

\newpage
\appendix
\input{appendix}

\end{document}

